---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# mantis

The mantis package generates interactive html reports that enable quick visual review of multiple
related time series. This can help with identification of temporal artefacts and unexpected
trends. Each group of time series is displayed together, with adjustable axes and tooltips showing the
individual dates and values.

The data frame containing the time series should be in long format, i.e.:

* one "timepoint" (date) column which will be used for the x-axes. This currently must be at a daily granularity, but values do not have to be consecutive.
* one "item" (character) column containing categorical values identifying distinct time series.
* one "value" (numeric) column containing the time series values which will be used for the y-axes.
* Optionally, a "group" (character) column containing categorical values which will be used to group the time series into different tabs on the report.

The `colspec` parameter maps the data frame columns to the above.

The reports are shareable and can contribute to forming a transparent record of the entire
analysis process. It is designed with electronic health records in mind, but can be used for any
types of time series.

## Installation

```{r, eval = FALSE}
# install direct from source
# install.packages("remotes")
remotes::install_github("phuongquan/mantis")
```

## Usage

```{r}
library(mantis)

# this example data frame contains numbers of antibiotic prescriptions in long format, 
#   plus a column for grouping the output:
data("example_prescription_numbers")

head(example_prescription_numbers)

```

```{r, eval = FALSE}
# create a report in the current directory
mantis_report(
  df = example_prescription_numbers,
  colspec = colspec(
    timepoint_col = "PrescriptionDate",
    item_col = "Antibiotic",
    value_col = "NumberOfPrescriptions",
    group_col = "Location"
  )
)
```

## Acknowledgements

This work was supported by the National Institute for Health Research Health Protection Research Unit (NIHR HPRU) in Healthcare Associated Infections and Antimicrobial Resistance at the University of Oxford in partnership with Public Health England (PHE) (NIHR200915), and by the NIHR Oxford Biomedical Research Centre.


## Contributing to this package

Please report any bugs or suggestions by opening a [github issue](https://github.com/phuongquan/mantis/issues).
